---
format: gfm
default-image-extension: ""
editor_options: 
  chunk_output_type: console
---

# BiasCorrector 

<!-- badges: start -->
```{r}
#| echo: false
#| message: false
#| results: asis

pkg <- desc::desc_get_field("Package")
cat_var <- paste(
  badger::badge_lifecycle(),
  badger::badge_cran_release(pkg = pkg),
  gsub("summary", "worst", badger::badge_cran_checks(pkg = pkg)),
  badger::badge_cran_download(pkg = pkg, type = "grand-total", color = "blue"),
  badger::badge_cran_download(pkg = pkg, type = "last-month", color = "blue"),
  gsub("netlify\\.com", "netlify.app", badger::badge_dependencies(pkg = pkg)),
  badger::badge_github_actions(action = utils::URLencode("R CMD Check via {tic}")),
  badger::badge_github_actions(action = "lint"),
  badger::badge_github_actions(action = "test-coverage"),
  badger::badge_codecov(ref = desc::desc_get_urls()),
  badger::badge_doi("10.1002/ijc.33681", "yellow"),
  sep = "\n"
)
cat_var |> cat()
```
<!-- badges: end -->

`BiasCorrector` is published in *'BiasCorrector: fast and accurate correction of all types of experimental biases in quantitative DNA methylation data derived by different technologies' (2021)* in the *International Journal of Cancer* (DOI: [https://onlinelibrary.wiley.com/doi/10.1002/ijc.33681](https://doi.org/10.1002/ijc.33681)).

`BiasCorrector` is the user friendly implementation of the algorithms described by Moskalev et. al in their research article *'Correction of PCR-bias in quantitative DNA methylation studies by means of cubic polynomial regression'*, published 2011 in *Nucleic acids research, Oxford University Press* (DOI: [https://doi.org/10.1093/nar/gkr213](https://doi.org/10.1093/nar/gkr213)).  

## Installation  

### Using R 

- Make sure, you have R installed on your system:  
    
    + https://cran.r-project.org/  

- Then open your development environment and install this R package:

#### CRAN version

You can install `BiasCorrector` simply with via R's `install.packages` interface:

```{r}
#| eval: false
install.packages("BiasCorrector")
```

#### Development version

If you want to use the latest development version, you can install the github version of `BiasCorrector` with:

```{r}
#| eval: false
install.packages("remotes")
remotes::install_github("kapsner/BiasCorrector")
```

- To start BiasCorrector, just run the following command in R. A browser tab should open displaying BiasCorrector. Alternatively you can type the URL "localhost:3838/" in your browser.

```{r}
#| eval: false
library(BiasCorrector)
launch_app()
```

### Using Docker

To simplify installation an deployment of `BiasCorrector` you can clone this repository and build your own docker image. Make sure, you have Docker and docker-compose installed on your system.

#### Build Docker Image Manually

```bash
# clone the repository
git clone https://github.com/kapsner/BiasCorrector

# go to the docker subfolder
cd BiasCorrector/docker/

# run the build script
./build_image.sh

# when the building is finished, just start the container by running
docker-compose -f docker-compose.local.yml up -d
```

#### Using a Remote Docker Image

```bash
# clone the repository
git clone https://github.com/kapsner/BiasCorrector

# go to the docker subfolder
cd BiasCorrector/docker/

# start the Docker container
docker-compose -f docker-compose.remote.yml up -d
```

Type the URL "localhost:3838/" in your browser and start working with `BiasCorrector`. 

## rBiasCorrection

`BiasCorrector` depends on the `rBiasCorrection` R-package, which is the implementation of the core functionality to correct measurement biases in DNA methylation analyses. `BiasCorrector` brings this functionality to a user-friendly shiny web application.  
`rBiasCorrection` is available at [https://github.com/kapsner/rBiasCorrection](https://github.com/kapsner/rBiasCorrection). 

## Video Tutorial 

A video tutorial describing the workflow of how to use `BiasCorrector` in order to correct measurement bias in DNA methylation data is available [on youtube](https://youtu.be/xOf8uDbUrms). 

## Demo Version

A demo version of `BiasCorrector` is available [here](https://biascorrector.diz.uk-erlangen.de/).

## Frequently Asked Questions 

More detailed information on how to use the backend-package `rBiasCorrection` can be found in its [vignette](https://cran.r-project.org/web/packages/rBiasCorrection/vignettes/rBiasCorrection_howto.html). The FAQs can be found [here](https://github.com/kapsner/rBiasCorrection/blob/master/FAQ.md).

## Citation 

L.A. Kapsner, M.G. Zavgorodnij, S.P. Majorova, A. Hotz‐Wagenblatt, O.V. Kolychev, I.N. Lebedev, J.D. Hoheisel, A. Hartmann, A. Bauer, S. Mate, H. Prokosch, F. Haller, and E.A. Moskalev, BiasCorrector: fast and accurate correction of all types of experimental biases in quantitative DNA methylation data derived by different technologies, Int. J. Cancer. (2021) ijc.33681. doi:[10.1002/ijc.33681](https://onlinelibrary.wiley.com/doi/10.1002/ijc.33681).

```bibtex
@article{kapsner2021,
  title = {{{BiasCorrector}}: Fast and Accurate Correction of All Types of Experimental Biases in Quantitative {{DNA}} Methylation Data Derived by Different Technologies},
  author = {Kapsner, Lorenz A. and Zavgorodnij, Mikhail G. and Majorova, Svetlana P. and Hotz-Wagenblatt, Agnes and Kolychev, Oleg V. and Lebedev, Igor N. and Hoheisel, J{\"o}rg D. and Hartmann, Arndt and Bauer, Andrea and Mate, Sebastian and Prokosch, Hans-Ulrich and Haller, Florian and Moskalev, Evgeny A.},
  year = {2021},
  month = may,
  pages = {ijc.33681},
  issn = {0020-7136, 1097-0215},
  doi = {10.1002/ijc.33681},
  journal = {International Journal of Cancer},
  language = {en}
}
```

## More Infos

- Original work by Moskalev et al.: https://doi.org/10.1093/nar/gkr213
- about Shiny: https://www.rstudio.com/products/shiny/
- RStudio and Shiny are trademarks of RStudio, Inc.
- about Docker: https://www.docker.com/
